A Meta-analysis of Longevity Estimates of Mosquito Vectors of Disease: Additional Methods

### 2020-11-21 writeUp/supplementary_materials_methods.tex
updated graph titles

The R file \verb|s_clean_data.R| cleans the data, makes the above amendments and outputs the data in a form amenable for estimation using Stan \citep{carpenter2016stan}.

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mapGenusCropped.pdf}}
	\caption{\textbf{The location and number of separate time-series for the MRR studies included in this analysis.} The area of each circle indicates the number of time-series at each study site. The colour shows the genus of the mosquitoes in each study.}
	\label{fig:mrr_lifetimes_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lc}
		\toprule
		\textbf{Data characteristic} & \textbf{Count} \\
		\midrule
		Species & 33 \\
		Genera & 3 \\
		\textit{Anopheles} & 94 \\
		\textit{Aedes} & 91 \\
		\textit{Culex} & 47 \\
		Female-only MRR & 179 \\
		Male-only MRR & 35 \\
		Mixed sex MRR & 18 \\
		Pre-release blood-feeding only MRR & 71 \\
		Pre-release sugar-feeding only MRR & 41 \\
		Pre-release both blood- and sugar-feeding MRR & 4 \\
		Pre-release neither blood- and sugar-feeding MRR & 116 \\
		MRR time-series & 232 \\
		\bottomrule
	\end{tabular}%
	\caption{\textbf{Summary of variables across all MRR time-series used in hierarchical analyses.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
	\label{tab:mrr_aggregateData}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.5in}{-0.5in}%
		\begin{tabularx}{1.25\textwidth}{l|ccccc}
			\toprule
			\textbf{Data characteristic} & \textbf{Min} & \textbf{Mean} & \textbf{Median} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Study duration, days & 6.0   & 11.8  & 10.0  & 71.0    & 4.7 \\
			Number of days on which collections took place & 6.0   & 10.1  & 9.0   & 47    & 9.1 \\
			Number of separate release days & 1.0     & 1.9   & 1.0     & 23    & 3.0 \\
			Number released & 66    & 4,929  & 1,297  & 86,200 & 12,043 \\
			Number recaptured & 2     & 163   & 63    & 4,090  & 399 \\
			Recapture percentage & 0\% & 8.6\% & 5.2\% & 57.1\% & 10.04\% \\
			Age at release, days & 0   & 1.8   & 0   & 13    & 2.8 \\
			\bottomrule
		\end{tabularx}%
		\caption{\textbf{Summary of data from individual MRR time-series used in hierarchical analyses.}}
		\label{tab:mrr_IndividualData}%
	\end{adjustwidth}
\end{table}%

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
	\caption{\textbf{The number of MRR time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
	\label{fig:mrr_sexReleasesOverTime}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_exampleMRRSeries.pdf}}
	\caption{\textbf{Example MRR time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments, there was a single release of marked mosquitoes on day zero.}
	\label{fig:mrr_exampleMRRSeries}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_individualTimeSeries_priors.pdf}}
	\caption{\textbf{The prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis for the MRR studies.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
	\label{fig:mrr_individualTimeSeries_priors}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_methodDiagramHierarchy.pdf}}
	\caption{\textbf{A representation of the hierarchical Bayesian model, for the constant mortality model case when considering the species-level grouping.} Here `pdf' indicates `probability density function', and $\lambda_i$, $\psi_i$, and $\kappa_i$ represent the force of mortality for the exponential distribution, daily recapture probability and over-dispersion parameter of the negative binomial distribution for time-series $i$ in the dataset; $\mu_i(t)= \psi_i e^{-\lambda_i t}$ is the mean number of mosquitoes recaptured on day $t$.}
	\label{fig:mrr_methodDiagramHierarchy}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespan_priors.pdf}}
	\caption{\textbf{Samples from the prior predictive distribution of mean lifespan across the six models of mosquito mortality introduced in Table \ref{tab:mrr_survivalDescription}.} In all cases, 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
	\label{fig:mrr_meanLifespanPrior}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_priors_PsiKappa.pdf}}
	\caption{\textbf{Samples from the priors for daily recapture probability ($\psi$) and over-dispersion parameter ($\kappa$).} In both panels, 10,000 samples were generated using the following hierarchical priors: for the individual time series parameters, $\psi_i \sim \text{exp}(r)$ (with $\psi$ constrained to be less than 0.1 - encompassing the upper limit of day one recapture fractions observed in most data series) and $\kappa_i \sim \text{exp}(p)$; for the top-level parameters: $r\sim \text{gamma}(100, 2)$ and $p\sim \text{log-normal}(1.5, 1)$.}\label{fig:mrr_PsiKappaPriors}
\end{figure}

@@

\begin{table}
	\footnotesize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.2\textwidth}{l|c|X}
			\textbf{Survival function} & \textbf{time-series-level priors} & \textbf{Group-level priors}\\
			\midrule
			Exponential & $c\sim \mathcal{N}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-2.32,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-3,1)$\\
			\midrule
			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,1.1)$\\
			\midrule
			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-4.8,1.75)\text{,}\; \mu_\beta\sim N(-4,2)$\\
			\midrule
			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
			\midrule
			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-2.8,1)\text{,}\; \mu_\beta\sim N(-3,1)\text{,}\; \mu_s\sim N(-3,1)$\\
			\midrule
			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3.2,1)\text{,}\; \mu_\beta\sim N(-3.2,1)\text{,}\; \mu_s\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
			\bottomrule
	\end{tabularx}}\caption{\textbf{The priors used on parameters of each different survival model.} For the exponential model, the `group-level' priors were the same for the genus and `overall' models that were also estimated. The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data. The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:mrr_priors}
\end{table}

@@

{%
	\includegraphics[page=1,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc1}
	\par

{%
	\includegraphics[page=2,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc2}
	\par
}

{%
	\includegraphics[page=3,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{The numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc3}
	\par
}

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=0.8\textwidth]{./Figure_files/dissection_timeSeries.pdf}}
	\caption{\textbf{The numbers of published studies that estimate gonotrophic age by Polovodova's dissection method across the study period.}}\label{fig:dissection_timeSeries}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/dissection_map.pdf}}
	\caption{\textbf{The geographic location of each of the Polovodova datasets included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lll}
		\toprule
		\textbf{Genus} & \textbf{Species} & \textbf{Frequency} \\
		\midrule
		\textit{Anopheles} & \textit{gambiae s.l.} & 19 \\
		\textit{Culex} & \textit{quinquefasciatus} & 12 \\
		\textit{Anopheles} & \textit{maculipennis} & 11 \\
		\textit{Anopheles} & \textit{farauti s.l.} & 10 \\
		\textit{Anopheles} & \textit{sergentii} & 8 \\
		\textit{Aedes} & \textit{polynesiensis} & 8 \\
		\textit{Culex} & \textit{pipiens} & 6 \\
		\textit{Anopheles} & \textit{culicifacies} & 5 \\
		\textit{Anopheles} & \textit{darlingi} & 5 \\
		\textit{Anopheles} & \textit{quadrimaculatus} & 5 \\
		\textit{Anopheles} & \textit{stephensi} & 4 \\
		\textit{Anopheles} & \textit{melas} & 4 \\
		\textit{Culex} & \textit{annulirostris} & 4 \\
		\textit{Aedes} & \textit{aegypti} & 3 \\
		\textit{Aedes} & \textit{samoanus} & 3 \\
		\textit{Anopheles} & \textit{minimus} & 3 \\
		\textit{Anopheles} & \textit{rivulorum} & 3 \\
		\textit{Culex} & \textit{thalassius} & 3 \\
		\textit{Mansonia} & \textit{uniformis} & 3 \\
		\textit{Anopheles} & \textit{subpictus} & 2 \\
		\textit{Aedes} & \textit{sollicitans} & 2 \\
		\textit{Aedes} & \textit{vexans} & 2 \\
		\textit{Anopheles} & \textit{bellator} & 2 \\
		\textit{Anopheles} & \textit{cruzii} & 2 \\
		\textit{Culex} & \textit{tritaeniorhynchus} & 2 \\
		\bottomrule
		\textit{\textbf{Anopheles}} &       & \textbf{83} \\
		\textit{\textbf{Culex}} &       & \textbf{27} \\
		\textit{\textbf{Aedes}} &       & \textbf{18} \\
		\textit{\textbf{Mansonia}} &       & \textbf{3} \\
		\bottomrule
		\textbf{Total} &  & \textbf{131} \\
	\end{tabular}%
	\caption{\textbf{The numbers of Polovodova dissection series for each species or genus included in the overall dataset.} Note, these counts represent numbers of series analysed by the hierarchical models -- that is, after aggregating the individual series as described in Section \ref{sec:dissection_dissectionData}.}\label{tab:dissection_speciesNumbers}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.3in}{-0.3in}%
		\begin{tabularx}{1.1\textwidth}{l|ccccc}
			\toprule
			\textbf{Variable} & \textbf{Min} & \textbf{Median} & \textbf{Mean} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Min age of captures, gonotrophic cycles & 0  & 0  & 0.02  & 1.00  & 0.12 \\
			Median age of captures, gonotrophic cycles & 0  & 1.00  & 0.63  & 2.00  & 0.62 \\
			Mean age of captures, gonotrophic cycles & 0.17  & 0.90  & 1.00  & 3.03  & 0.58 \\
			Max age of captures, gonotrophic cycles & 1.00  & 5.00  & 5.77  & 13.00 & 2.94 \\
			Number captured & 100 & 565 & 1,317 & 14,012 & 1,885 \\
			\bottomrule
		\end{tabularx}%
	\end{adjustwidth}
	\caption{\textbf{Summaries of the characteristics of the 131 individual Polovodova dissection data series that are included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
\end{table}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data1.png}}
	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each cas,e the count for all ages has been normalised by the nulliparous count. In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData1}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data2.png}}
	\caption{\textbf{The normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case, the count for all ages has been normalised by the nulliparous count. In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData2}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data3.png}}
	\caption{\textbf{The normalised physiological age series for seven species in the database.} Each different coloured line represents an individual series. In each case, the count for all ages has been normalised by the nulliparous count. In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData3}
\end{figure}

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
	\caption{\textbf{The priors for mean lifespan used to analyse the Polovodova dissection data for each of the hierarchical models.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}. In all cases, the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
\end{figure}

>>>>

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_mapGenusCropped.pdf}}
	\caption{\textbf{MRR: location of studies included in the analysis.} The area of each circle indicates the number of time-series at each study site. The colour shows the genus of the mosquitoes in each study.}
	\label{fig:mrr_lifetimes_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lc}
		\toprule
		\textbf{Data characteristic} & \textbf{Count} \\
		\midrule
		Species & 33 \\
		Genera & 3 \\
		\textit{Anopheles} & 94 \\
		\textit{Aedes} & 91 \\
		\textit{Culex} & 47 \\
		Female-only MRR & 179 \\
		Male-only MRR & 35 \\
		Mixed sex MRR & 18 \\
		Pre-release blood-feeding only MRR & 71 \\
		Pre-release sugar-feeding only MRR & 41 \\
		Pre-release both blood- and sugar-feeding MRR & 4 \\
		Pre-release neither blood- and sugar-feeding MRR & 116 \\
		MRR time-series & 232 \\
		\bottomrule
	\end{tabular}%
	\caption{\textbf{MRR: summary of time-series used in hierarchical analyses.} For the `Species' and `Genera' rows, the counts represent the number of distinct entries in the database; otherwise the counts represent the number of time series.}
	\label{tab:mrr_aggregateData}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.5in}{-0.5in}%
		\begin{tabularx}{1.25\textwidth}{l|ccccc}
			\toprule
			\textbf{Data characteristic} & \textbf{Min} & \textbf{Mean} & \textbf{Median} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Study duration, days & 6.0   & 11.8  & 10.0  & 71.0    & 4.7 \\
			Number of days on which collections took place & 6.0   & 10.1  & 9.0   & 47    & 9.1 \\
			Number of separate release days & 1.0     & 1.9   & 1.0     & 23    & 3.0 \\
			Number released & 66    & 4,929  & 1,297  & 86,200 & 12,043 \\
			Number recaptured & 2     & 163   & 63    & 4,090  & 399 \\
			Recapture percentage & 0\% & 8.6\% & 5.2\% & 57.1\% & 10.04\% \\
			Age at release, days & 0   & 1.8   & 0   & 13    & 2.8 \\
			\bottomrule
		\end{tabularx}%
		\caption{\textbf{MRR: summary of data from individual time-series used in hierarchical analyses.}}
		\label{tab:mrr_IndividualData}%
	\end{adjustwidth}
\end{table}%

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_sexReleasesOverTime.pdf}}
	\caption{\textbf{MRR: time-series by date and sex of released mosquitoes.} The numbers represent by-decade totals of the 232 time-series we used in the main analysis.}
	\label{fig:mrr_sexReleasesOverTime}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_exampleMRRSeries.pdf}}
	\caption{\textbf{MRR: time-series for three different studies from the \cite{guerra2014global} database.} In all the above experiments, there was a single release of marked mosquitoes on day zero.}
	\label{fig:mrr_exampleMRRSeries}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/mrr_individualTimeSeries_priors.pdf}}
	\caption{\textbf{MRR: prior probability distributions used for lifespan (A), recapture probability (B) and dispersal parameter (C) in the individual time-series analysis.} Note: the prior for the recapture probability was truncated at $\psi=0.1$ since most MRR day one recapture fractions were significantly below this threshold.}
	\label{fig:mrr_individualTimeSeries_priors}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_methodDiagramHierarchy.pdf}}
	\caption{\textbf{MRR: structure of hierarchical Bayesian model.} This diagram represents the structure of the model for the constant mortality model case with the species-level grouping. Here `pdf' indicates `probability density function', and $\lambda_i$, $\psi_i$, and $\kappa_i$ represent the force of mortality for the exponential distribution, daily recapture probability and over-dispersion parameter of the negative binomial distribution for time-series $i$ in the dataset; $\mu_i(t)= \psi_i e^{-\lambda_i t}$ is the mean number of mosquitoes recaptured on day $t$.}
	\label{fig:mrr_methodDiagramHierarchy}
\end{figure}

@@

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_lifespan_priors.pdf}}
	\caption{\textbf{MRR: prior predictive distribution of mean lifespan across the six models of mosquito mortality.} In all cases, 10,000 samples were generated using the priors indicated in Table \ref{tab:mrr_priors}, resulting in distributions with a mean close to 10 days.}
	\label{fig:mrr_meanLifespanPrior}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/mrr_priors_PsiKappa.pdf}}
	\caption{\textbf{MRR: samples from the priors for daily recapture probability ($\psi$) and over-dispersion parameter ($\kappa$).} In both panels, 10,000 samples were generated using the following hierarchical priors: for the individual time series parameters, $\psi_i \sim \text{exp}(r)$ (with $\psi$ constrained to be less than 0.1 - encompassing the upper limit of day one recapture fractions observed in most data series) and $\kappa_i \sim \text{exp}(p)$; for the top-level parameters: $r\sim \text{gamma}(100, 2)$ and $p\sim \text{log-normal}(1.5, 1)$.}\label{fig:mrr_PsiKappaPriors}
\end{figure}

@@

\begin{table}
	\footnotesize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.2\textwidth}{l|c|X}
			\textbf{Survival function} & \textbf{time-series-level priors} & \textbf{Group-level priors}\\
			\midrule
			Exponential & $c\sim \mathcal{N}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-2.32,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-3,1)$\\
			\midrule
			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,1.1)$\\
			\midrule
			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-4.8,1.75)\text{,}\; \mu_\beta\sim N(-4,2)$\\
			\midrule
			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
			\midrule
			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-2.8,1)\text{,}\; \mu_\beta\sim N(-3,1)\text{,}\; \mu_s\sim N(-3,1)$\\
			\midrule
			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3.2,1)\text{,}\; \mu_\beta\sim N(-3.2,1)\text{,}\; \mu_s\sim N(-3,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
			\bottomrule
	\end{tabularx}}\caption{\textbf{MRR: priors used on parameters of each survival model.} For the exponential model, the `group-level' priors were the same for the genus and `overall' models that were also estimated. The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data. The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:mrr_priors}
\end{table}

@@

{%
	\includegraphics[page=1,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{\textbf{MRR: posterior predictive distribution versus actual data.} The figure shows the numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc1}
	\par

{%
	\includegraphics[page=2,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{\textbf{MRR: posterior predictive distribution versus actual data.} The figure shows the numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc2}
	\par
}

{%
	\includegraphics[page=3,scale=.6]{./Figure_files/mrr_ppcs_all}
	\captionof{figure}{\textbf{MRR: posterior predictive distribution versus actual data.} The figure shows the numbers of mosquitoes recaptured (black lines) versus the 95\% central posterior interval of the posterior predictive distribution (blue shading) for a selection of the time series. For the rest of the posterior predictive checks for the MRR models, see the file referenced in the text.}
	\label{fig:mrr_ppc3}
	\par
}

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=0.8\textwidth]{./Figure_files/dissection_timeSeries.pdf}}
	\caption{\textbf{Polovodova dissection: numbers of published studies that estimate gonotrophic age by this method.}}\label{fig:dissection_timeSeries}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.25\textwidth]{./Figure_files/dissection_map.pdf}}
	\caption{\textbf{Polovodova dissection: the geographic location of each of the datasets included in the meta-analysis.} The area of the bubbles indicates the number of unique data series available.}\label{fig:dissection_map}
\end{figure}

\begin{table}[htbp]
	\centering
	\begin{tabular}{lll}
		\toprule
		\textbf{Genus} & \textbf{Species} & \textbf{Frequency} \\
		\midrule
		\textit{Anopheles} & \textit{gambiae s.l.} & 19 \\
		\textit{Culex} & \textit{quinquefasciatus} & 12 \\
		\textit{Anopheles} & \textit{maculipennis} & 11 \\
		\textit{Anopheles} & \textit{farauti s.l.} & 10 \\
		\textit{Anopheles} & \textit{sergentii} & 8 \\
		\textit{Aedes} & \textit{polynesiensis} & 8 \\
		\textit{Culex} & \textit{pipiens} & 6 \\
		\textit{Anopheles} & \textit{culicifacies} & 5 \\
		\textit{Anopheles} & \textit{darlingi} & 5 \\
		\textit{Anopheles} & \textit{quadrimaculatus} & 5 \\
		\textit{Anopheles} & \textit{stephensi} & 4 \\
		\textit{Anopheles} & \textit{melas} & 4 \\
		\textit{Culex} & \textit{annulirostris} & 4 \\
		\textit{Aedes} & \textit{aegypti} & 3 \\
		\textit{Aedes} & \textit{samoanus} & 3 \\
		\textit{Anopheles} & \textit{minimus} & 3 \\
		\textit{Anopheles} & \textit{rivulorum} & 3 \\
		\textit{Culex} & \textit{thalassius} & 3 \\
		\textit{Mansonia} & \textit{uniformis} & 3 \\
		\textit{Anopheles} & \textit{subpictus} & 2 \\
		\textit{Aedes} & \textit{sollicitans} & 2 \\
		\textit{Aedes} & \textit{vexans} & 2 \\
		\textit{Anopheles} & \textit{bellator} & 2 \\
		\textit{Anopheles} & \textit{cruzii} & 2 \\
		\textit{Culex} & \textit{tritaeniorhynchus} & 2 \\
		\bottomrule
		\textit{\textbf{Anopheles}} &       & \textbf{83} \\
		\textit{\textbf{Culex}} &       & \textbf{27} \\
		\textit{\textbf{Aedes}} &       & \textbf{18} \\
		\textit{\textbf{Mansonia}} &       & \textbf{3} \\
		\bottomrule
		\textbf{Total} &  & \textbf{131} \\
	\end{tabular}%
	\caption{\textbf{Polovodova dissection: numbers of dissection series included in the overall dataset.} Note, these counts represent numbers of series analysed by the hierarchical models -- that is, after aggregating the individual series as described in Section \ref{sec:dissection_dissectionData}.}\label{tab:dissection_speciesNumbers}%
\end{table}%

\begin{table}[htbp]
	\centering
	\footnotesize
	\begin{adjustwidth}{-0.3in}{-0.3in}%
		\begin{tabularx}{1.1\textwidth}{l|ccccc}
			\toprule
			\textbf{Variable} & \textbf{Min} & \textbf{Median} & \textbf{Mean} & \textbf{Max} & \textbf{Standard deviation} \\
			\midrule
			Min age of captures, gonotrophic cycles & 0  & 0  & 0.02  & 1.00  & 0.12 \\
			Median age of captures, gonotrophic cycles & 0  & 1.00  & 0.63  & 2.00  & 0.62 \\
			Mean age of captures, gonotrophic cycles & 0.17  & 0.90  & 1.00  & 3.03  & 0.58 \\
			Max age of captures, gonotrophic cycles & 1.00  & 5.00  & 5.77  & 13.00 & 2.94 \\
			Number captured & 100 & 565 & 1,317 & 14,012 & 1,885 \\
			\bottomrule
		\end{tabularx}%
	\end{adjustwidth}
	\caption{\textbf{Polovodova dissection: summaries of the data series included in this analysis.} Note when series are censored (see Section \ref{sec:dissection_censored}) we assume that all mosquitoes dissected over the threshold are of this age when calculating summary statistics.}\label{tab:dissection_summaryStats}
\end{table}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data1.png}}
	\caption{\textbf{Polovodova dissection: normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each cas,e the count for all ages has been normalised by the nulliparous count. In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData1}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data2.png}}
	\caption{\textbf{Polovodova dissection: normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case, the count for all ages has been normalised by the nulliparous count. In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData2}
\end{figure}

\begin{figure}[ht]
	\centerline{\includegraphics[width=1.3\textwidth]{./Figure_files/dissection_parity_data3.png}}
	\caption{\textbf{Polovodova dissection: normalised physiological age series for nine species in the database.} Each different coloured line represents an individual series. In each case, the count for all ages has been normalised by the nulliparous count. In all cases, we do not include any data for censored observations (see Section \ref{sec:dissection_censored}).}\label{fig:dissection_exampleData3}
\end{figure}

@@

\begin{figure}[ht]
	\centerline{\includegraphics[width=1\textwidth]{./Figure_files/dissection_lifespanPriorsParity.pdf}}
	\caption{\textbf{Polovodova dissection: prior predictive mean lifespan for each mortality model.} The distributional form of the priors is given in Table \ref{tab:dissection_priors}. In all cases, the plots show data for 10,000 samples from the relevant prior distribution.}\label{fig:dissection_lifespanPriorsParity}
\end{figure}

\begin{table}
	\footnotesize
	\noindent\makebox[\textwidth]{%
		\begin{tabularx}{1.2\textwidth}{l|c|X}
			\textbf{Survival function} & \textbf{Physiological age series-level priors} & \textbf{Group-level priors}\\
			\midrule
			Exponential & $\lambda\sim \textnormal{log-normal}(\mu_\lambda,\sigma)$ &$\mu_\lambda \sim N(-1.2,1)\text{,}\; \sigma\sim \textnormal{log-normal}(-2,1)$\\
			\midrule
			Gompertz & $\alpha,\beta\sim$ log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-1.5,1)\text{,}\; \mu_\beta\sim N(-2.5,0.5)$\\
			\midrule
			Weibull & $\alpha,(\beta-1)\sim$  log-normal$(\mu_{\alpha|\beta},0.2)$ &$\mu_\alpha\sim N(-2.5,1)\text{,}\; \mu_\beta\sim N(-4,0.5)$\\
			\midrule
			Gompertz-Makeham & $\alpha,\beta,c\sim$ log-normal$(\mu_{\alpha|\beta|c},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-3,0.4)\text{,}\; \mu_c\sim N(-4.5,0.5)$\\
			\midrule
			Logistic & $\alpha,\beta,s\sim$  log-normal$(\mu_{\alpha|\beta|s},0.2)$ &$\mu_\alpha\sim N(-1.4,1)\text{,}\; \mu_\beta\sim N(-2,1)\text{,}\; \mu_s\sim N(-3,1)$\\
			\midrule
			Logistic-Makeham & $\alpha,\beta,s,c\sim$ log-normal$(\mu_{\alpha|\beta|s|c},0.2)$ &$\mu_\alpha\sim N(-3,1)\text{,}\; \mu_\beta\sim N(-1.3,1)\text{,}\; \mu_s\sim N(-2,0.5)\text{,}\; \mu_c\sim N(-4,0.5)$\\
			\bottomrule
	\end{tabularx}}\caption{\textbf{Polovodova dissection: priors used on parameters of each different survival model.} For the exponential model the `group-level' priors were the same for the genus and `overall' models that were also estimated. The exponential model was the only model that was simple enough to allow the scale parameter of the log-normal ($\sigma$) to be estimated by the data. The notation $\alpha,\beta\sim \text{log-normal}(\mu_{\alpha|\beta},0.2)$ means that $\alpha$ and $\beta$ were assigned independent log-normal priors with location parameters $\mu_\alpha$ and $\mu_\beta$ respectively, and a scale parameter of 0.2 in both cases.}\label{tab:dissection_priors}
\end{table}
